---
title: "Milestone 2"
author: "Alyssa Monda"
date: "3/28/2017"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#Downloaded packages 
library(phyloseq)
library(dada2)
library(ggplot2)
library(biom)
library(devtools)
library(knitr)
library(mixOmics)
library(tidyverse)
library(vegan)
library(mixOmics)
library(readxl)
library(tableone)
library(ReporteRs)
library(magrittr)
library(plyr)
```

##Change in Plans
Originally I had described using a different dataset to investigate the relationship between butyrate producing taxa (BPT) of the gut microbiome and inflammatory conditions. After beginning the data wrangling process I discovered the variables within the dataset did not match up in a way that made answering my question possible. The majority of the samples were from the oral microbiome and very few individuals had a positive value for an inflammatory disease as well as a gut sample, most of the time they only had oral samples on these individuals. While looking for a new dataset I stumbled upon the data from American Gut and realized they had 3 variables relating to physical activity. 
Although this may seem like a significant change in plans, my purpose for investigating the aforementioned data was to assess whether a relationship existed between BPT and chronic conditions, because there is an increase in BPT seen in animal models observing exercise and the gut microbiome. Little is known about the mechanisms involved in the relationship between exercise and reduction of inflammation. I hypothesize changes in the gut microbiome as one of the possible pathways for the anti-inflammatory effects of moderate intensity, routine, exercise. I knew this task would involve investing more time into cleaning and analyzing this data, its significance to my research interests made the additional effort worth it. To my knowledge there are only 3 studies published investigating the relationship between exercise and the gut microbiome. These studies have small sample sizes, and the methods are of varying quality. With the current plan to propose a longitudinal study to characterize the gut microbiome after participation in a routine, aerobic, exercise intervention for my dissertation, the data from this study will be a wonderful addition to my future proposal.

##Study Aims
Aim 1: Characterize the alpha/beta diversity of the gut microbiome as it relates to frequency and location of exercise. 
  
Aim 2: If a relationship exists, investigate the associations between exercise and butyrate producing taxa.  

##Import Meta Data
Import Meta Data as DataFrame
```{r}
library(readr)
Meta <- read_delim("~/Desktop/ag_1k_fecal.txt", 
    "\t", escape_double = FALSE, col_types = cols(AGE_YEARS = col_number(), 
        BMI = col_number(), `ENA-BASE-COUNT` = col_number(), 
        HEIGHT_CM = col_number(), VIOSCREEN_FIBER = col_number(),
        VIOSCREEN_LOW_FAT_DAIRY_SERVING = col_number(), 
        VIOSCREEN_LYSINE = col_number(), 
        VIOSCREEN_MANNITOL = col_number(), 
        VIOSCREEN_SFA100 = col_double(), #capric acid
        VIOSCREEN_SFA40 = col_double(),#SFA4:0 Butyric Acid
        VIOSCREEN_SFA170 = col_double(), #margaric acid
        VIOSCREEN_VEG5_DAY = col_number(), 
        VIOSCREEN_VEGETABLE_SERVINGS = col_number(), 
        chao1_1k = col_double()), trim_ws = TRUE)
View(Meta)

#Import Tree File ###Not Working 
#tree_file <-("~/Desktop/97_otus.tree")
#treefilename = system.file("~/Desktop/97_otus.tree", "biom-tree.phy",  package="phyloseq")

#What are parsing failures?
```

##Analysis of MetaData
It is important to clean the metadata prior to merging it with the data from the biom file. Therefore I will now recode, clean, and analyze the demographics of the sample. Before I can start I must recategorize the variables and assign values to groups I am interested in. 

```{r}

#Recode labels and levels for current variables to clean up data
#Age
Meta$Age<-Meta$"AGE_YEARS"

#BMI- coded as wanted check implausible values after recoding other variables

#Sex 
Meta$Sex<-Meta$"SEX"
Meta$Sex[Meta$Sex=="male"] <- "Male"
Meta$Sex[Meta$Sex=="female"] <- "Female"
Meta$Sex[Meta$Sex=="other"] <- "Other"
Meta$Sex[Meta$Sex=="Unknown"] <- "NA"
Meta$Sex[Meta$Sex=="unspecified"] <- "NA"
Meta$Sex

#Race 
#What I want: If Race="White" then Race="White", If African American then Race="African American", else="Other"
Meta$Race<- Meta$"RACE"
Meta$Race[Meta$Race=="Caucasian"] <- "White"
Meta$Race[Meta$Race=="African American"] <- "Black"
Meta$Race[Meta$Race=="Hispanic"] <- "Hispanic"
Meta$Race[Meta$Race=="Asian or Pacific Islander"] <- "Other"
Meta$Race[Meta$Race=="Other"] <- "Other"
Meta$Race[Meta$Race=="Unspecified"] <- "NA"
Meta$Race[Meta$Race=="Unknown"] <- "NA"
Meta$Race

#Education 
Education<- "LEVEL_OF_EDUCATION"
Meta$Education<- Meta$"LEVEL_OF_EDUCATION"
Meta$Education[Meta$Education=="Did not complete high school"] <- "< High School Degree"
Meta$Education[Meta$Education=="High School or GED equivalent degree"] <- "High School Degree"
Meta$Education[Meta$Education=="Some college or technical school"] <- "High School Degree"
Meta$Education[Meta$Education=="Associate's degree"] <- "Undergraduate degree"
Meta$Education[Meta$Education=="Bachelor's degree"] <- "Undergraduate degree"
Meta$Education[Meta$Education=="Some graduate school or professional"] <- "Undergraduate degree"
Meta$Education[Meta$Education=="Graduate or Professional degree"] <- "Graduate or Professional degree"
Meta$Education[Meta$Education=="Uknown"] <- "NA"
Meta$Education[Meta$Education=="Unspecified"] <- "NA"

#Do I want to include country of birth, or economic region?

#Antibiotics 
Meta$Antibiotics<-Meta$"ANTIBIOTIC_HISTORY"
Meta$Antibiotics[Meta$Antibiotics=="6 months"] <- "6 Months"
Meta$Antibiotics[Meta$Antibiotics=="Uknown"] <- "> 1 year"
Meta$Antibiotics[Meta$Antibiotics=="Uknown"] <- "NA"
Meta$Antibiotics[Meta$Antibiotics=="Unspecified"] <- "NA"

#Probiotics
Meta$Probiotics<- Meta$"PROBIOTIC_FREQUENCY"
Meta$Probiotics[Meta$Probiotics=="Unknown"] <- "NA"
Meta$Probiotics[Meta$Probiotics=="Unspecified"] <- "NA"

#ExFreq 
Meta$ExFreq <-Meta$"EXERCISE_FREQUENCY"
Meta$ExFreq[Meta$ExFreq=="Unknown"] <- "NA"
Meta$ExFreq[Meta$ExFreq=="Unspecified"] <- "NA"

#ExLoc 
Meta$ExLoc<- Meta$"EXERCISE_LOCATION"
Meta$ExLoc[Meta$ExLoc=="Depends on the season"] <- "Both"
Meta$ExLoc[Meta$ExLoc=="Unknown"] <- "NA"
Meta$ExLoc[Meta$ExLoc=="Unspecified"] <- "NA"

#PoolFreq 
Meta$PoolFreq<- Meta$"POOL_FREQUENCY"
Meta$PoolFreq[Meta$PoolFreq=="Unknown"] <- "NA"
Meta$PoolFreq[Meta$PoolFreq=="Unspecified"] <- "NA"

#Diet
Meta$Diet<- Meta$"DIET_TYPE"
Meta$Diet[Meta$Diet=="Unknown"] <- "NA"
Meta$Diet[Meta$Diet=="Unspecified"] <- "NA"

#Healthy = No chronic Illness
Meta$Healthy<-Meta$"SUBSET_HEALTHY"
Meta$Healthy[Meta$Healthy=="True"] <- "Healthy"
Meta$Healthy[Meta$Healthy=="False"] <- "Chronic Illness"
Meta$Healthy[Meta$Healthy=="true"] <- "Healthy"
Meta$Healthy[Meta$Healthy=="false"] <- "Chronic Illness"
Meta$Healthy[Meta$Healthy=="Uknown"] <- "NA"
Meta$Healthy[Meta$Healthy=="Unspecified"] <- "NA"

#Recode variables for chronic illnesses to 0-Do not have 1-Dx by provider NA-unknown

Meta$Cancer=Meta$"CANCER"
Meta$Cancer[Meta$Cancer=="Diagnosed by a medical professional doctor, physician assistant"] <- "1"
Meta$Cancer[Meta$Cancer=="Diagnosed by a medical professional (doctor, physician assistant)"] <- "1"
Meta$Cancer[Meta$Cancer=="Diagnosed by an alternative medicine practitioner"] <- "1"
Meta$Cancer[Meta$Cancer=="Self-diagnosed"] <- "0"
Meta$Cancer[Meta$Cancer=="I do not have this condition"] <- "0"
Meta$Cancer[Meta$Cancer=="Unknown"] <- "NA"
Meta$Cancer[Meta$Cancer=="Unspecified"] <- "NA"
Meta$Cancer[Meta$Cancer=="no_data"] <- "NA"



Meta$Kidney=Meta$"KIDNEY_DISEASE"
Meta$Kidney[Meta$Kidney=="Diagnosed by a medical professional doctor, physician assistant"] <- "1"
Meta$Kidney[Meta$Kidney=="Diagnosed by a medical professional (doctor, physician assistant)"] <- "1"
Meta$Kidney[Meta$Kidney=="Diagnosed by an alternative medicine practitioner"] <- "1"
Meta$Kidney[Meta$Kidney=="Self-diagnosed"] <- "0"
Meta$Kidney[Meta$Kidney=="I do not have this condition"] <- "0"
Meta$Kidney[Meta$Kidney=="Unknown"] <- "NA"
Meta$Kidney[Meta$Kidney=="Unspecified"] <- "NA"
Meta$Kidney[Meta$Kidney=="no_data"] <- "NA"


Meta$DM=Meta$"DIABETES"
Meta$DM[Meta$DM=="Diagnosed by a medical professional doctor, physician assistant"] <- "1"
Meta$DM[Meta$DM=="Diagnosed by a medical professional (doctor, physician assistant)"] <- "1"
Meta$DM[Meta$DM=="Diagnosed by an alternative medicine practitioner"] <- "1"
Meta$DM[Meta$DM=="Self-diagnosed"] <- "0"
Meta$DM[Meta$DM=="I do not have this condition"] <- "0"
Meta$DM[Meta$DM=="Unknown"] <- "NA"
Meta$DM[Meta$DM=="Unspecified"] <- "NA"
Meta$DM[Meta$DM=="no_data"] <- "NA"


Meta$Lung=Meta$"LUNG_DISEASE"
Meta$Lung[Meta$Lung=="Diagnosed by a medical professional doctor, physician assistant"] <- "1"
Meta$Lung[Meta$Lung=="Diagnosed by a medical professional (doctor, physician assistant)"] <- "1"
Meta$Lung[Meta$Lung=="Diagnosed by an alternative medicine practitioner"] <- "1"
Meta$Lung[Meta$Lung=="Self-diagnosed"] <- "0"
Meta$Lung[Meta$Lung=="I do not have this condition"] <- "0"
Meta$Lung[Meta$Lung=="Unknown"] <- "NA"
Meta$Lung[Meta$Lung=="Unspecified"] <- "NA"
Meta$Lung[Meta$Lung=="no_data"] <- "NA"


Meta$ASD[Meta$ASD=="Diagnosed by a medical professional doctor, physician assistant"] <- "1"
Meta$ASD[Meta$ASD=="Diagnosed by a medical professional (doctor, physician assistant)"] <- "1"
Meta$ASD[Meta$ASD=="Diagnosed by an alternative medicine practitioner"] <- "1"
Meta$ASD[Meta$ASD=="Self-diagnosed"] <- "0"
Meta$ASD[Meta$ASD=="I do not have this condition"] <- "0"
Meta$ASD[Meta$ASD=="Unknown"] <- "NA"
Meta$ASD[Meta$ASD=="Unspecified"] <- "NA"
Meta$ASD[Meta$ASD=="no_data"] <- "NA"


Meta$IBS[Meta$IBS=="Diagnosed by a medical professional doctor, physician assistant"] <- "1"
Meta$IBS[Meta$IBS=="Diagnosed by a medical professional (doctor, physician assistant)"] <- "1"
Meta$IBS[Meta$IBS=="Diagnosed by an alternative medicine practitioner"] <- "1"
Meta$IBS[Meta$IBS=="Self-diagnosed"] <- "0"
Meta$IBS[Meta$IBS=="I do not have this condition"] <- "0"
Meta$IBS[Meta$IBS=="Unknown"] <- "NA"
Meta$IBS[Meta$IBS=="Unspecified"] <- "NA"
Meta$IBS[Meta$IBS=="no_data"] <- "NA"

 
Meta$IBD[Meta$IBD=="Diagnosed by a medical professional doctor, physician assistant"] <- "1"
Meta$IBD[Meta$IBD=="Diagnosed by a medical professional (doctor, physician assistant)"] <- "1"
Meta$IBD[Meta$IBD=="Diagnosed by an alternative medicine practitioner"] <- "1"
Meta$IBD[Meta$IBD=="Self-diagnosed"] <- "0"
Meta$IBD[Meta$IBD=="I do not have this condition"] <- "0"
Meta$IBD[Meta$IBD=="Unknown"] <- "NA"
Meta$IBD[Meta$IBD=="Unspecified"] <- "NA"
Meta$IBD[Meta$IBD=="no_data"] <- "NA"


Meta$Liver=Meta$"LIVER_DISEASE"
Meta$Liver[Meta$Liver=="Diagnosed by a medical professional doctor, physician assistant"] <- "1"
Meta$Liver[Meta$Liver=="Diagnosed by a medical professional (doctor, physician assistant)"] <- "1"
Meta$Liver[Meta$Liver=="Diagnosed by an alternative medicine practitioner"] <- "1"
Meta$Liver[Meta$Liver=="Self-diagnosed"] <- "0"
Meta$Liver[Meta$Liver=="I do not have this condition"] <- "0"
Meta$Liver[Meta$Liver=="Unknown"] <- "NA"
Meta$Liver[Meta$Liver=="Unspecified"] <- "NA"
Meta$Liver[Meta$Liver=="no_data"] <- "NA"


Meta$CVD=Meta$"CARDIOVASCULAR_DISEASE"
Meta$CVD[Meta$CVD=="Diagnosed by a medical professional doctor, physician assistant"] <- "1"
Meta$CVD[Meta$CVD=="Diagnosed by a medical professional (doctor, physician assistant)"] <- "1"
Meta$CVD[Meta$CVD=="Diagnosed by an alternative medicine practitioner"] <- "1"
Meta$CVD[Meta$CVD=="Self-diagnosed"] <- "0"
Meta$CVD[Meta$CVD=="I do not have this condition"] <- "0"
Meta$CVD[Meta$CVD=="Unknown"] <- "NA"
Meta$CVD[Meta$CVD=="Unspecified"] <- "NA"
Meta$CVD[Meta$CVD=="no_data"] <- "NA"

 
Meta$AutoImm=Meta$"AUTOIMMUNE"
Meta$AutoImm[Meta$AutoImm=="Diagnosed by a medical professional doctor, physician assistant"] <- "1"
Meta$AutoImm[Meta$AutoImm=="Diagnosed by a medical professional (doctor, physician assistant)"] <- "1"
Meta$AutoImm[Meta$AutoImm=="Diagnosed by an alternative medicine practitioner"] <- "1"
Meta$AutoImm[Meta$AutoImm=="Self-diagnosed"] <- "0"
Meta$AutoImm[Meta$AutoImm=="I do not have this condition"] <- "0"
Meta$AutoImm[Meta$AutoImm=="Unknown"] <- "NA"
Meta$AutoImm[Meta$AutoImm=="Unspecified"] <- "NA"
Meta$AutoImm[Meta$AutoImm=="no_data"] <- "NA"


Meta$AD=Meta$"ALZHEIMERS"
Meta$AD[Meta$AD=="Diagnosed by a medical professional doctor, physician assistant"] <- "1"
Meta$AD[Meta$AD=="Diagnosed by a medical professional (doctor, physician assistant)"] <- "1"
Meta$AD[Meta$AD=="Diagnosed by an alternative medicine practitioner"] <- "1"
Meta$AD[Meta$AD=="Self-diagnosed"] <- "0"
Meta$AD[Meta$AD=="I do not have this condition"] <- "0"
Meta$AD[Meta$AD=="Unknown"] <- "NA"
Meta$AD[Meta$AD=="Unspecified"] <- "NA"
Meta$AD[Meta$AD=="no_data"] <- "NA"


Meta$Allergies=Meta$"SEASONAL_ALLERGIES"
Meta$Allergies[Meta$Allergies=="Diagnosed by a medical professional doctor, physician assistant"] <- "1"
Meta$Allergies[Meta$Allergies=="Diagnosed by a medical professional (doctor, physician assistant)"] <- "1"
Meta$Allergies[Meta$Allergies=="Diagnosed by an alternative medicine practitioner"] <- "1"
Meta$Allergies[Meta$Allergies=="Self-diagnosed"] <- "0"
Meta$Allergies[Meta$Allergies=="I do not have this condition"] <- "0"
Meta$Allergies[Meta$Allergies=="Unknown"] <- "NA"
Meta$Allergies[Meta$Allergies=="Unspecified"] <- "NA"
Meta$Allergies[Meta$Allergies=="no_data"] <- "NA"


Meta$Thyroid=Meta$"THYROID"
Meta$Thyroid[Meta$Thyroid=="Diagnosed by a medical professional doctor, physician assistant"] <- "1"
Meta$Thyroid[Meta$Thyroid=="Diagnosed by a medical professional (doctor, physician assistant)"] <- "1"
Meta$Thyroid[Meta$Thyroid=="Diagnosed by an alternative medicine practitioner"] <- "1"
Meta$Thyroid[Meta$Thyroid=="Self-diagnosed"] <- "0"
Meta$Thyroid[Meta$Thyroid=="I do not have this condition"] <- "0"
Meta$Thyroid[Meta$Thyroid=="Unknown"] <- "NA"
Meta$Thyroid[Meta$Thyroid=="Unspecified"] <- "NA"
Meta$Thyroid[Meta$Thyroid=="no_data"] <- "NA"


Meta$Skin=Meta$"SKIN_CONDITION"
Meta$Skin[Meta$Skin=="Diagnosed by a medical professional doctor, physician assistant"] <- "1"
Meta$Skin[Meta$Skin=="Diagnosed by a medical professional (doctor, physician assistant)"] <- "1"
Meta$Skin[Meta$Skin=="Diagnosed by an alternative medicine practitioner"] <- "1"
Meta$Skin[Meta$Skin=="Self-diagnosed"] <- "0"
Meta$Skin[Meta$Skin=="I do not have this condition"] <- "0"
Meta$Skin[Meta$Skin=="Unknown"] <- "NA"
Meta$Skin[Meta$Skin=="Unspecified"] <- "NA"
Meta$Skin[Meta$Skin=="no_data"] <- "NA"

```

Look for implausible values in numeric vasriables that need to be removed from the analysis, and observe distributions. 
```{r}
#Histogram for Age
hist(Meta$Age)
#Should probably remove kids from the analysis but I'm not sure how this would then affect the merge to the .biom file. Ask about this!*

#Histogram for BMI
hist(Meta$BMI)

#Remove implausible values BMI. Remove any value over 100.Seems mistakes are all in the thousands or higher. New max is 94.48. Need to find method for what is implausible. Obese Class III is >40. Best I have found is any score 3 standard deviations from the mean.<http://apps.who.int/bmi/index.jsp?introPage=intro_3.html> I can also use BMI_CAT for classifications. 
Meta$BMI[Meta$BMI > 100] <- NA

# Re-Do Histogram for BMI
hist(Meta$BMI)

#Scatterplot BMI*AGE
plot(Meta$BMI, Meta$Age)
```

Now I will create lists of variables to make analysis easier for the groupings I know I want to use. 
```{r}

#Create a list of all variables of interest
listVars <- c(Age="Age", AgeCat="AGE_CAT", Sex="Sex", Race="Race", BMI="BMI", Education="Education", Antibiotics="Antibiotics", Probiotics="Probiotics", ExFreq="ExFreq", ExLoc="ExLoc", PoolFreq="PoolFreq", Diet="Diet", Healthy="Healthy", Cancer="Cancer", Kidney="Cancer", DM="DM", Lung="Lung", CVD="CVD", AD="AD", Allergies="Allergies", AutoImm="AutoImm", ASD="ASD", IBD="IBD", IBS="IBS", Liver="Liver", Thyroid="Thyroid", Skin="Skin")

#Demographic Variables for Table 1
DemVars <- c(Age="Age", AgeCat="AGE_CAT", Sex="Sex", Race="Race", BMI="BMI", Education="Education", Healthy="Healthy")

#Variables to describe lifestyle
LifeVars<- c(Antibiotics="Antibiotics", Probiotics="Probiotics", ExLoc="ExLoc", PoolFreq="PoolFreq", Diet="Diet")

#Specific Illnesses of the sample
Chronic<- c(Cancer="Cancer", Kidney="Cancer", DM="DM", Lung="Lung", CVD="CVD", AD="AD", Allergies="Allergies", AutoImm="AutoImm", ASD="ASD", IBD="IBD", IBS="IBS", Liver="Liver", Thyroid="Thyroid", Skin="Skin")

#Categorical Variables
catVars <- c(Sex="Sex", Race="Race", Education="Education", Antibiotics="Antibiotics", Probiotics="Probiotics", ExFreq="ExFreq", ExLoc="ExLoc", PoolFreq="PoolFreq", Diet="Diet", Healthy="Healthy", Cancer="Cancer", Kidney="Cancer", DM="DM", Lung="Lung", CVD="CVD", AD="AD", Allergies="Allergies", AutoImm="AutoImm", ASD="ASD", IBD="IBD", IBS="IBS", Liver="Liver", Thyroid="Thyroid", Skin="Skin")

#Continuous Variables
contVars <- c(Age="Age", BMI="BMI")
```

Now that the data is reorganized and cleaned it is time to run an initial analysis to (1) see if recoding and grouping worked and (2) look at intitial analyais of the sample data. I will do so by group of self reported Exercise Frequency. 
```{r}
#Using "tableone" package

#Define categorical variables
CreateCatTable(vars= catVars, data=Meta, includeNA = FALSE, test = TRUE,testApprox = chisq.test, argsApprox = list(correct = TRUE),
  testExact = fisher.test, argsExact = list(workspace = 2 * 10^5),
  smd = TRUE)

#Total Population Table 1
table1 <- CreateTableOne(vars = listVars, data = MetaIdeal, factorVars = catVars)
table1
```


Create Table 1 for men and women and to compare their means and proportions to see if this is a more meaningful way to present the data. 
```{r}
# Create table by exercise frequency category 
#Define categorical variables

ExTable1<- CreateTableOne(vars=list2Vars, strata=ExFreq, data=D, factorVars=catVars, includeNA = FALSE,
  test = TRUE, testApprox = chisq.test, argsApprox = list(correct = TRUE),
  testExact = fisher.test, argsExact = list(workspace = 2 * 10^5),
  testNormal = oneway.test, argsNormal = list(var.equal = TRUE),
  testNonNormal = kruskal.test, argsNonNormal = list(NULL), smd = TRUE),
```

If I wanted to export table 1 to Microsoft Word I would run the following code. 
```{r}
ExTable1 <- print(ExTable1),

# The script
docx( ) %>% 
     addFlexTable(ExTable1 %>%
     FlexTable(header.cell.props = cellProperties( background.color = "#003366"),
               header.text.props = textBold( color = "white" ),
               add.rownames = TRUE ) %>%
               setZebraStyle( odd = "#DDDDDD", even = "#FFFFFF" ) ) %>%
     writeDoc(file = "~/Desktop/Extable1.docx"),
```
***Need to recode categories for variables, and modify code to delete test column and add a total. Figure out how to make continuous variables show mean and SD. 
Also figure out how p value was calculated. 

Below are explanations for real tables I would like to make. 

Figure 1: Demographic Information by Exercise Frequency
```{r}
#Age, BMI, Race, Gender, Education, Region or country of Residence, chronic condition


```


Figure 2: Lifestyle habits by Exercise Frequency
```{r}
#Diet, Exercise location, pool, antibiotic, probiotic 


```


Figure 3: Chronic Conditions by Exercise Frequency
```{r}
#Investigate Chronic Condition Frequencies
ExTable3<- CreateTableOne(vars=Chronic, strata=ExFreq, data=Meta, factorVars=catVars, includeNA = FALSE,
  test = TRUE, testApprox = chisq.test, argsApprox = list(correct = TRUE),
  testExact = fisher.test, argsExact = list(workspace = 2 * 10^5),
  testNormal = oneway.test, argsNormal = list(var.equal = TRUE),
  testNonNormal = kruskal.test, argsNonNormal = list(NULL), smd = TRUE),
ExTable3
```

##Import OTU file
Now that the metadata is cleaned I can upload the .biom file and merge the meta data. 
```{r}
#Data originally downloaded from <ftp://ftp.microbio.me/AmericanGut/latest>
#Make sure to pull the correct biom file to match the meta data down for analysis. 

# To read in original .biom file use command below 
file_path<- ("~/Desktop/ag_1k_fecal.biom")
dat <- import_biom(file_path)

#To fix error "input string 1 is invalid in this locale" run command "Sys.setlocale(locale="C")" in console and run command chunk again

#Make Phyloseq values for separate pieces of the biom file 
#Taxonomy
taxonomy <- tax_table(dat)
#OTU
OTU_table<- otu_table(dat)

#MetaData
#Create Phyloseq object for sample data so it can be merged
SampleData<- sample_data(Meta)
SD<-sample_data(Meta)

#Import the tree corresponding to this data ### Will Not work
#tree<- tree

#Merge OTU, Taxonomy, and metadata into one phyloseq object called "ps" that can be used for analysis
Sample_data <- (SD)
sam_cov <- as.data.frame(Sample_data)
rownames(sam_cov) <- sam_cov$"#SampleID"
sd <- sample_data(sam_cov)
sample_names(sd) #Previous steps fix error in naming of SampleIds
otus <- otu_table(OTU_table, taxa_are_rows = TRUE)
sample_names(otus)
tt <- tax_table(taxonomy)
ps <- phyloseq(otus, sd, tt) #Step to merge OTU,Taxonomy, and metadata 
ps #View object and make sure it has all 3 parts

```


##Microbiome Analysis
Time to start the fun part! To begin, I will run a few investigatory plots to look at the sample data. 

Figure 4:: Richness by Exercise Frequency 
```{r}
plot_richness(ps, ExFreq, color = NULL, shape = NULL,
               title = NULL, scales = "free_y", nrow = 1, shsi = NULL,
             measures = NULL, sortby = NULL)
```

General ordination plot (Need more info on which method to use for this type of plot)
```{r}
#Command for plot_ordination says to run this command prior to running the command
Ordination<- ordinate(ps, method = "DCA", distance = "bray", formula = NULL),

#Ordinaton plot
plot_ordination(ps, Ordination, type = "taxa", axes = 1:2,
  color ="EXERCISE_FREQUENCY", shape = NULL, label = NULL, title = NULL,
  justDF = FALSE),
```

Barplot 
```{r}
#"Melt Data" as reccommended by plot_bar command to turn phyloseq object into a large table in order to build ggplot2 graphics from microbiome data represented as phyloseq objects
#meltps<- psmelt(ps), #comand won't work, shuts down my computer or causes R to restart

#Barplot
plot_bar(ps, x="Taxa", y="Abundance", fill="EXERCISE_FREQUENCY",
title="Bar Plot", facet_grid=NULL),
```


Heatmap

```{r}
#Create a heatmap in ggplot using previously made data frame using psmelt command
ggplot(meltps, aes(x = EXERCISE_FREQUENCY, 
                y = OTU, 
                fill = Abundance)) 

#Hetmap using phyloseq
plot_heatmap(ps, method = NULL, distance = NULL,
  sample.label = EXERCISE_FREQUENCY, taxa.label = "Rank 5", trans = log_trans(4),
  max.label = 250, title = NULL, sample.order = NULL, taxa.order = NULL,
  first.sample = NULL, first.taxa = NULL)
```


Figure 4: Alpha Diversity

Figure 5: 

Figure 6: Beta Diversity By Exercise Frequency

Figure 7: Beta Diversity By Exercise Location

##Preliminary Synthesis/Discussion of Results

##Ancillary Questions
-How do the skewed demographics affect the data? This group is mostly cacuasian and has at least a Bachelor's degreee or has attended graduate school. 
-Should children be removed from the analysis?
-Would including birth method/ method of feeding received as an infant change any of the results?

##References
Put references for the exercise part for why I changed 


##GitHub Repository
Ther RMarkdown file used to create this document can be found in the repositor at <https://github.com/amonda/American-Gut>. The name of the file is "Milestone2.Rmd".