---
title: "Starting Over Data Wrangle 3_17"
author: "Alyssa Monda"
date: "3/17/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import Data
Assign file pathways as value names to use in the next R Code chunk. 
```{r}
rich_dense_biom<- ("~/Desktop/Big Data/American-Gut/data/AG/AG_even1k.biom")
treefilename<- ("~/Desktop/Big Data/American-Gut/data/AG/BeyondBacteria/AmericanGutMetaphlanWGS.taxonomy.tree")
meta<- ("~/Desktop/Big Data/American-Gut/data/AG/AG_even1k.txt")
```

Use Import_Biom command to import the biom file, and tanxonomy tree. 
```{r}
import_biom(rich_dense_biom, parseFunction=parse_taxonomy_greengenes)
#import_biom(rich_dense_biom, treefilename, parseFunction=parse_taxonomy_greengenes) tree file name should be included but resulted in an error and failed to import?
```

Store import as myData file to use this data object in downstream data manipulations and analysis. 

```{r}
myData = import_biom(rich_dense_biom, parseFunction=parse_taxonomy_greengenes)
myData
```
Did it work?No.
```{r}
plot_bar(myData, fill="Genus")
```
Import metadata
```{r}
library(readxl)
Ag_MetaData1k_Trimmed <- read_excel("~/Desktop/Big Data/American-Gut/data/AG/Ag_MetaData1k-Trimmed.xlsx")
View(Ag_MetaData1k_Trimmed)
meta<-(Ag_MetaData1k_Trimmed)
```

#Split apart and merge data into proper format
```{r}
data(myData)
#tree = phy_tree("~/Desktop/Big Data/American-Gut/data/AG/BeyondBacteria/AmericanGutMetaphlanWGS.taxonomy.tree")
tax  = tax_table(myData)
otu  = otu_table(myData)
sam  = sample_data(meta)
otutax = phyloseq(otu, tax)
otutax
```

Now use merge_phyloseq to build a file in the proper format so it includes the information listed above. 
```{r}
MD2 = merge_phyloseq(otutax, sam)
```






