---
title: "Starting Over Data Wrangle 3_17"
author: "Alyssa Monda"
date: "3/17/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(phyloseq)
library(dada2)
library(ggplot2)
```

## Import Data
Assign file pathways as value names to use in the next R Code chunk. 
```{r}
rich_dense_biom<- ("~/Desktop/Big Data/American-Gut/data/AG/AG_even1k.biom")
treefilename<- ("~/Desktop/Big Data/American-Gut/data/AG/BeyondBacteria/AmericanGutMetaphlanWGS.taxonomy.tree")
meta<- ("~/Desktop/Big Data/American-Gut/data/AG/AG_even1k.txt")
```

Use Import_Biom command to import the biom file, and tanxonomy tree. 
```{r}
import_biom(rich_dense_biom, parseFunction=parse_taxonomy_greengenes)
#import_biom(rich_dense_biom, treefilename, parseFunction=parse_taxonomy_greengenes) tree file name should be included but resulted in an error and failed to import?
```

Store import as myData file to use this data object in downstream data manipulations and analysis. 

```{r}
myData = import_biom(rich_dense_biom, parseFunction=parse_taxonomy_greengenes)
myData
```
Did it work?No.
```{r}
plot_bar(myData, fill="Genus")
```

Import metadata From Excel Instead b/c the other did not import as data but as values

Export Values to csv file. 
```{r}
write.csv(otu, "~/Desktop/Big Data/American-Gut/data/AG/Otu1k.xlsx")

write.csv(tax, "~/Desktop/Big Data/American-Gut/data/AG/Tax1k.xlsx")
```


```{r}
library(readxl)
Ag_MetaData1k_Trimmed <- read_excel("~/Desktop/Big Data/American-Gut/data/AG/Ag_MetaData1k-Trimmed.xlsx")
View(Ag_MetaData1k_Trimmed)
meta<-(Ag_MetaData1k_Trimmed)

#OTU
OTU1k <- read_excel("~/Desktop/Big Data/American-Gut/data/AG/Ag_MetaData1k-Trimmed.xlsx")
View(Ag_MetaData1k_Trimmed)
meta<-(Ag_MetaData1k_Trimmed)

#Taxonomy
Taxonomy1k <- read_excel("~/Desktop/Big Data/American-Gut/data/AG/Ag_MetaData1k-Trimmed.xlsx")
View(Ag_MetaData1k_Trimmed)
meta<-(Ag_MetaData1k_Trimmed)
```

#Split apart and merge data into proper format
```{r}
data(myData)
#tree = phy_tree("~/Desktop/Big Data/American-Gut/data/AG/BeyondBacteria/AmericanGutMetaphlanWGS.taxonomy.tree")
tax  = tax_table(myData)
otu  = otu_table(myData)
sam  = sample_data(meta)
otutax = phyloseq(otu, tax)
otutax
```

Now use merge_phyloseq to build a file in the proper format so it includes the information listed above. 
```{r}
MD2 = merge_phyloseq(otutax, sam)
```



Code from Hertzberg.... 
I was getting the following error when I went to create the phyloseq object:

“Error in validObject(.Object) : invalid class “phyloseq” object: 
 Component sample names do not match.
 Try sample_names()”

I figured that it must have something to do with the sample data matrix that I was importing, and eventually figured out that it had to have “rownames” that were the same as the otu_table rownames. The following is my bit of code:

```{r}
#Origina
#library(readr)
#sample_data <- read_csv("~/Desktop/MYSEQFILES/sample_data.csv")
#sam_cov <- as.data.frame(sample_data)
#rownames(sam_cov) <- sam_cov$ID
#sd <- sample_data(sam_cov)
#sample_names(sd)
#otus <- otu_table(seqtab.nochim, taxa_are_rows = FALSE)
#sample_names(otus)
#tt <- tax_table(taxa)
#ps <- phyloseq(otus, sd, tt)
#Ps


#My rewrite for big data set but it still results in error, won't write sd
library(readr)
Sample_data <- (Meta)
sam_cov <- as.data.frame(sample_data)
rownames(sam_cov) <- sam_cov()$#SampleID
sd <- sample_data(sam_cov)
sample_names(sd)
otus <- otu_table(goods, taxa_are_rows = TRUE)
sample_names(otus)
tt <- tax_table(taxonomy)
ps <- phyloseq(otus, sd, tt)
Ps
```


